Faithful Variable Screening for High-Dimensional Convex Regression

### 2014-07-19 tex/scam/additive_faithful.tex
edits through 3.2

\begin{remark}
  We assume twice differentiability in
  Theorems~\ref{thm:convex_faithful} to simplify the proof. 
  We believe
  this smoothness condition is not necessary because every non-smooth
  convex function can be approximated arbitrarily well by a smooth
  one.  
\end{remark}

\begin{remark} 
It remains difficult to find conditions under which the opposite direction of additive faithfulness holds, that is, if $f$ does not depend on coordinate $k$, then $f_k^*$ will be zero in the additive approximation. As a conceptual example, suppose that $f$ is only a function of $X_1, X_2$, and that $(X_1, X_2, X_3)$ follows a degenerate 3-dimensional distribution where $X_3 = f(X_1, X_2) - f^*(X_1) - f^*_2(X_2)$, that is, we let $X_3$ exactly capture the additive approximation error. In this case, the best additive approximation of $f$ would have a component $f^*_3(X_3) = X_3$ even though $f$ does not depend on $X_3$. 
\end{remark}

It is possible to get a similar result for distributions with unbounded support, by using a limit condition $\lim_{|x_k| \rightarrow \infty} \frac{\partial p(\mathbf{x}_{-k} \given x_k)}{\partial x_k} = 0$ instead of the boundary condition. Such a limit condition is too strong however, it is not obeyed by the multivariate Gaussian distribution. \\

In fact, not all convex functions are additively faithful under certain multivariate Gaussian distributions.

\begin{example}
\label{examp:gaussian_counterexample}
Consider a quadratic form $f( \mathbf{x}) = \mathbf{x}^\tran H \mathbf{x} + c^\tran \mathbf{x}$ and a Gaussian distribution $X \sim N(0, \Sigma)$ where $\Sigma_{jj} = 1$ for all $j$.

Then we have a closed form for the additive approximation, which we prove in [TODO:appendix reference].
\begin{itemize}
\item If $f$ does not depend on $x_j$, then $f^*_j(x_j) = 0$.
\item If $f$ depends on $x_j$, then
\begin{equation}
f^*_j(x_j) = H_j^\tran \Sigma_j x_j^2 + c_j x_j
\end{equation}
where $H_j, \Sigma_j$ are the $j$-th row of $H_j, \Sigma_j$ respectively. 
\end{itemize}
Let $H = [1, 2; 2,5]$ and let $\Sigma = [1, -\alpha; -\alpha, 1]$. It is easy to check then that if $\alpha=0.5$, then $f^*_1 = 0$ and additive faithfulness is violated, and if $\alpha > 0.5$, then $f^*_1$ is a concave function.
\end{example}

>>>>

\begin{remark}
  We assume twice differentiability in
  Theorems~\ref{thm:convex_faithful} to simplify the proof.  We
  expect, however, that this this smoothness condition is not
  necessary---every convex function can be approximated arbitrarily
  well by a smooth convex function.
\end{remark}

\begin{remark} 
  We have not found natural conditions under which the opposite
  direction of additive faithfulness holds---conditions implying that if $f$ does not
  depend on coordinate $k$, then $f_k^*$ will be zero in the additive
  approximation.  Suppose, for example, that $f$ is only a
  function of $X_1, X_2$, and that $(X_1, X_2, X_3)$ follows a
  degenerate 3-dimensional distribution where $X_3 = f(X_1, X_2) -
  f^*(X_1) - f^*_2(X_2)$.  In this case $X_3$ exactly captures the
  additive approximation error.  The best additive
  approximation of $f$ would have a component $f^*_3(x_3) = x_3$ even
  though $f$ does not depend on $x_3$.
\end{remark}

\begin{remark}
It is possible to get a similar result for distributions with
unbounded support, by using a limit condition $\lim_{|x_k| \rightarrow
  \infty} \frac{\partial p(\mathbf{\scriptstyle x}_{-k} \given x_k)}{\partial x_k}
= 0$.  Such a limit condition is too
strong however, as it is not obeyed by the multivariate Gaussian
distribution. 
\end{remark}

The following example shows that not all convex functions are
additively faithful under multivariate Gaussian distributions.

\begin{example}
\label{examp:gaussian_counterexample}
Consider a quadratic form $f( \mathbf{x}) = \mathbf{x}^\tran H \mathbf{x} + c^\tran \mathbf{x}$ and a Gaussian distribution $X \sim N(0, \Sigma)$ where $\Sigma_{jj} = 1$ for all $j$.
As we show in the appendix, the additive approximation has the
following closed form.
\begin{enumerate}
\item If $f$ does not depend on $x_j$, then $f^*_j(x_j) = 0$.
\item If $f$ depends on $x_j$, then
\begin{equation}
f^*_j(x_j) = H_j^\tran \Sigma_j x_j^2 + c_j x_j
\end{equation}
where $H_j, \Sigma_j$ are the $j$-th rows of $H_j, \Sigma_j$, respectively. 
\end{enumerate}
Let $H = \begin{pmatrix}1 & 2 \\ 2 & 5\end{pmatrix}$ and let $\Sigma
= \begin{pmatrix}1 & -\alpha \\ -\alpha & 1\end{pmatrix}$. 
It can be checked that if if $\alpha=\frac{1}{2}$, then $f^*_1 = 0$
and additive faithfulness is violated.  Moreover, if $\alpha > \frac{1}{2}$, then $f^*_1$ is a concave function.
\end{example}
